---
title: "Advanced Graphics in R"
subtitle: "ggplot2: A Layered Grammar of Graphics"
author: "Iowa State University"
date: "`r Sys.Date()`"
output:
  ioslides_presentation:
    transition: default
    widescreen: true
css:
  styles.css
---

```{r, message = FALSE, echo = FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align='center', cache=TRUE, tidy=FALSE, warning = FALSE, message = FALSE)
```

## BP Oil Spill

<div align = "center">
  <img src = "satelitepicBPoilspill.png">
</div>

## Data Sets

1. NOAA Data
    * National Oceanic and Atmospheric Administration
    * Temperature and Salinity Data in Gulf of Mexico
    * Measured using Floats, Gliders and Boats

2. US Fisheries and Wildlife Data
    * Animal Sightings on the Gulf Coast
    * Birds, Turtles and Mammals
    * Status: Oil Covered or Not

Both data sets have geographic coordinates for ever observation

## Loading NOAA Data

```{r}
x <- ls()
noaa <- "http://www.public.iastate.edu/~hofmann/looking-at-data/data/noaa.rdata"
if (!file.exists("noaa.rdata")) download.file(noaa, "noaa.rdata")
load("noaa.rdata")
setdiff(ls(), x)
```

## Floats Data

```{r}
str(floats)
```

## Floats

```{r}
library(ggplot2)
qplot(Longitude, Latitude, colour = callSign, data = floats) + 
  coord_map()
```

## A note on `qplot()` versus `ggplot()`

* `qplot()` makes many assumptions to save you some typing.
* Although it's often useful for quick, interactive, exploratory data analysis; eventually you'll want the full flexibility of `ggplot()`.
* `qplot()` assumes variable names are _aesthetics_ (a mapping from data to visual elements), but with `ggplot()`, you have to specify them.

```{r, eval = FALSE}
ggplot(data = floats, aes(x = Longitude, y = Latitude, colour = callSign)) +
  geom_point() + coord_map()
```

* [RStudio's ggplot2 cheatsheet](http://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf) provides a great visual of this difference.

## Gliders Data

```{r}
str(gliders)
```

## Gliders

```{r}
ggplot(data = gliders, aes(x = Longitude, y = Latitude, colour = callSign)) + 
  geom_point() + coord_map()
```

## Boats Data

```{r}
str(boats)
```

## Boats

```{r}
ggplot(data = boats, aes(x = Longitude, y = Latitude, colour = callSign)) + 
  geom_point() + coord_map()
```

## Provide some context

* All of this data comes from the same geographic region.
* It'd be nice if we could _overlay_ this information onto a map!
* This is where ggplot2's layering idea becomes useful.

---

```{r}
states <- map_data("state")
map_outline <- ggplot() +
  geom_path(data = states, aes(x = long, y = lat, group = group)) + 
  xlim(c(-91, -80)) + ylim(c(22,32)) + coord_map()
map_outline
```

---

```{r}
map_floats <- map_outline +
  geom_point(data = floats, aes(x = Longitude, y = Latitude, colour = callSign))
map_floats
```

----

```{r}
rig # location of BP Oil rig
map_floats +
  geom_point(data = rig, aes(x, y), shape = "x", size = 5) + 
  geom_text(data = rig, aes(x, y), label = "BP Oil rig", 
            size = 5, hjust = -0.1)
```

## Your Turn

* Use your ggplot2 skills to explore any (or all!) of the floats/gliders/boats data. Be creative!!

# How _exactly_ does layering work?

## A simpler example

```{r}
p <- ggplot(data = diamonds, aes(x = carat, y = price)) +
  geom_point() +           # layer 1
  geom_smooth(method = lm) # layer 2
p
```

## Plot-level things

```{r}
# scales
p <- p + scale_x_log10() + scale_y_log10()  
p
```

---

```{r}
# coordinate system
p + coord_polar()
```

---

```{r}
# facets
p + facet_grid(. ~ cut)
```

---

<div align = "center">
  <a href = "http://www.voovarb.com/ggplot2Intro/ggplot_structure.png">
    <img src = "ggplot_structure.png" width = "1000" height = "1200">
  </a>
</div>

---

* Each `geom_*` function is really a layer with certain defaults for statistic, position, and, well, geometry.

```{r}
args(geom_point)
args(geom_smooth)
args(geom_bar)
```

## Defaults aren't always right!!

```{r, out.extra='style="float:left"', fig.show = "hold", fig.width=3.3}
base <- ggplot(diamonds, aes(x = cut, fill = clarity)) + theme(legend.position = "none")
base + geom_bar()                                   # bad!!!
base + geom_bar(position = "fill")                  # better
base + geom_bar(position = "dodge") + 
  theme(legend.position = "bottom")                 # better
```

## Your Turn

* Read in the animal.csv data

```{r}
animals <- read.csv("http://heike.github.io/rwrks/02-r-graphics/data/animal.csv")
```

* Plot the location of animal sightings on a map of the region
* On this plot, try to color points by class of animal and/or
status of animal
* Advanced: Could we indicate time somehow?

