---
title: "Reading Files"
author: "Eric Hare, Andee Kaplan, Samantha Tyner"
date: "May 18, 2016"
output: 
    revealjs::revealjs_presentation:
        transition: fade
        theme: white
        highlight: zenburn
        css: ../../css/slides.css
---

## Data in Excel

- Formats xls and csv - what's the difference?
- File extensions xls and xlsx are proprietary Excel formats, binary files
- csv is an extension for Comma Separated Value files. They are text files - directly readable.
- Example: Gas prices in midwest since 1994

## Gas Prices in the Midwest

```{r}
midwest <- read.csv("http://heike.github.io/rwrks/03-r-format/data/midwest.csv")
head(midwest)
```

## Gas Prices

```{r}
str(midwest)
```

There is clearly some work to be done...

## `read.csv` vs. `read.table`

`read.csv` is really just a wrapper for `read.table` with certain parameters set:

```{r}
read.csv
```

So to properly read in this data, it's probably best to use `read.table` directly

```{r, eval=FALSE}
?read.table
```

## Your Turn

Have a look at the parameters of read.table (?read.table) to solve the following problems:

1. Read the first two lines of the file into an object called `midwest_names`
2. Read everything EXCEPT the first two lines into an object called `midwest_data`

```{r, echo=FALSE}
midwest_names <- read.table("http://heike.github.io/rwrks/03-r-format/data/midwest.csv", nrows = 2, sep = ",", stringsAsFactors = FALSE)
midwest_data <- read.table("http://heike.github.io/rwrks/03-r-format/data/midwest.csv", skip = 2, sep = ",", stringsAsFactors = FALSE)
```

## Quick Lesson in Data Cleaning

```{r, warning=FALSE, message=FALSE}
values <- c(midwest_data$V3, midwest_data$V5, midwest_data$V7, 
            midwest_data$V9, midwest_data$V11)
dates <- c(paste(midwest_data$V1, midwest_data$V2, sep = "-"), 
           paste(midwest_data$V1, midwest_data$V4, sep = "-"),
           paste(midwest_data$V1, midwest_data$V6, sep = "-"),
           paste(midwest_data$V1, midwest_data$V8, sep = "-"),
           paste(midwest_data$V1, midwest_data$V10, sep = "-"))

dates <- dates[!is.na(values)]
values <- values[!is.na(values)]

library(lubridate)
dates <- ymd(dates)

midwest_gas <- data.frame(date = dates, price = values)
midwest_gas <- midwest_gas[with(midwest_gas, order(date)), ]
```

## Now we can make a plot!

```{r, warning=FALSE, message=FALSE, fig.width=7, fig.height=4}
library(ggplot2)
qplot(date, price, data = midwest_gas, geom = "line")
```

## Reading Excel Data

Download the midwest.xls file to your current working directory (`getwd()`)

```{r, eval=FALSE}
library(readxl)

midwest2 <- read_excel("midwest.xls")

head(midwest2)
```

```{r, echo=FALSE}
library(readxl)

midwest2 <- read_excel("../data/midwest.xls")

head(midwest2)
```

## `foreign` Package

- Other file formats can be read using the functions from package \texttt{foreign}
- SPSS: `read.spss`
- SAS: `read.xport`
- Minitab: `read.mtp`
- Systat: `read.systat`

## Your Turn

The NHANES (National Health and Nutrition Survey) publishes data in the SAS xport format:

http://wwwn.cdc.gov/Nchs/Nhanes/Search/nhanes13_14.aspx

1. Scroll to the bottom, choose one of the datasets (Demographics, Examination, etc.). Download the Data file (XPT)
2. Use `read.xport()` to load the file into R
