---
title: "Intro to dplyr"
author: "Eric Hare, Andee Kaplan, Samantha Tyner"
date: "May 18, 2016"
output: revealjs::revealjs_presentation
---

## Baseball Data

- The `plyr` package contains the data set `baseball`
- seasonal batting statistics of all major league players (through 2007)
 
```{r}
data(baseball, package = "plyr")
help(baseball, package = "plyr")
head(baseball)
```

## Baseball Data

- We would like to create career summary statistics for each player
- Plan: subset on a player, and compute statistics

```{r}
ss <- subset(baseball, id == "sosasa01")
head(ss)
mean(ss$h / ss$ab)
```

We need an **automatic** way to calculate this.

## `for` loops

- Idea: repeat the same (set of) statement(s) for each element of an index set
- Setup: 
    - Introduce counter variable (sometimes named `i`)
    - Reserve space for results
- Generic Code:

```{r, eval=FALSE}
result <- rep(NA, length(indexset))
for(i in indexset){
  ... some statements ...
  result[i] <- ...
}
```
    
## `for` loops for Baseball

- Index set: player id
- Setup: 

```{r}
players <- unique(baseball$id)
n <- length(players)

ba <- rep(NA, n)

for(i in 1:n) {
  career <- subset(baseball, id == players[i])
  ba[i] <- with(career, mean(h / ab, na.rm = TRUE))
}

summary(ba)
```

## `for` loops for Baseball

- Index set: player id
- `i = 0`: 

```{r}
players <- unique(baseball$id)
n <- length(players)

ba <- rep(NA, n)

head(ba)
```

## `for` loops for Baseball

- Index set: player id
- `i = 1`: 

```{r}
players <- unique(baseball$id)

ba <- rep(NA, length(players))

for(i in 1:1) {
  career <- subset(baseball, id == players[i])
  ba[i] <- with(career, mean(h / ab, na.rm = TRUE))
}

head(ba)
```

## `for` loops for Baseball

- Index set: player id
- `i = 2`: 

```{r}
players <- unique(baseball$id)

ba <- rep(NA, length(players))

for(i in 1:2) {
  career <- subset(baseball, id == players[i])
  ba[i] <- with(career, mean(h / ab, na.rm = TRUE))
}

head(ba)
```

## Your Turn

- MLB rules for the greatest all-time hitters are that players have to have played at least 1000 games with at least as many at-bats in order to be considered
- Extend the for loop above to collect the additional information
- Introduce and collect data for two new variables: `games` and `atbats`

## How did it go? What was difficult?

- Household chores (declaring variables, setting values each time) distract from real work
- Indices are error-prone
- Loops often result in slow code because R can compute quantities using entire vectors in an optimized way

## Summarise

A special function: `summarise` or `summarize`

```{r, message=FALSE}
library(dplyr)

baseball <- read.csv("http://heike.github.io/rwrks/03-r-format/data/baseball.csv")

summarise(baseball,
          ba = mean(h / ab, na.rm = TRUE),
          games = sum(g, na.rm = TRUE),
          hr = sum(hr, na.rm = TRUE),
          ab = sum(ab, na.rm = TRUE))
```

## More Summarise

```{r}
summarise(subset(baseball, id == "sosasa01"), 
          ba = mean(h / ab, na.rm = TRUE),
          games = sum(g, na.rm = TRUE),
          hr = sum(hr, na.rm = TRUE),
          ab = sum(ab, na.rm = TRUE))
```

## `dplyr` + `Summarise`

- A powerful combination to create summary statistics

```{r}
careers <- summarise(group_by(baseball, id),
                 ba = mean(h / ab, na.rm = TRUE),
                 games = sum(g, na.rm = TRUE),
                 homeruns = sum(hr, na.rm = TRUE),
                 atbats = sum(ab, na.rm = TRUE))

head(careers)
```

## Your Turn

- Find some summary statistics for each of the teams (variable `team`)
    - How many different (`unique`) players has the team had?
    - What was the team's first/last season?
- Challenge: Find the number of players on each team over time. Does the number change?
