---
title: "Plotting Map Data"
author: "Eric Hare, Andee Kaplan, Samantha Tyner"
date: "May 16, 2016"
output: revealjs::revealjs_presentation
---

```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
```

## States Data

To make a map, let's load up the states data and take a look:

```{r}
states <- map_data("state")

head(states)
```

## Basic Map Data

What needs to be in the data set in order to plot a basic map?
  
- Need latitude/longitude points for all map boundaries
- Need to know which boundary group all lat/long points belong
- Need to know the order to connect points within each group

## Data for Building Basic State Map

Our *states* data has all necessary information

## A Basic (Rather Hideous) Map

A bunch of latitude longitude points...

```{r, fig.height=4, fig.width=7}
qplot(long, lat, geom = "point", data = states)
```

## A Bit Better of a Map

... that are connected with lines in a very specific order.

```{r, fig.height=4, fig.width=7}
qplot(long, lat, geom = "path", data = states, group = group) + 
    coord_map()
```

## Polygon instead of Path

```{r, fig.height=4, fig.width=7}
qplot(long, lat, geom = "polygon", data = states, group = group) + 
    coord_map()
```

## Incorporating Information About States

We want to incorporate additional information into the plot:

- Add other geographic information by adding geometric layers to the plot
- Add non-geopgraphic information by altering the fill color for each state
    - Use `geom = "polygon"` to treat states as solid shapes to add color
    - Incorporate numeric information using color shade or intensity
    - Incorporate categorical informaion using color hue
    
## Categorical Information Using Hue

If a categorical variable is assigned as the fill color then `qplot` will assign different hues for each category. Let's load in a state regions dataset:

```{r}
statereg <- read.csv("http://heike.github.io/rwrks/02-r-graphics/data/statereg.csv")

head(statereg)
```

## Joining Data

We need to `join` or `merge` our original states data with this new information on the regions. We can use the `left_join` function to do so (more about this later):

```{r}
states.class.map <- left_join(states, statereg, by = c("region" = "State"))
head(states.class.map)
```

## Plotting the Result

```{r, fig.height=4, fig.width=7}
qplot(long, lat, geom = "polygon", data = states.class.map, 
      group = group, fill = StateGroups, colour = I("black")) + 
    coord_map() 
```

## Numerical Information Using Shade and Intensity

To show how was can add numerical information to map plots we will use the BRFSS data

- Behavioral Risk Factor Surveillance System
- 2008 telephone survey run by the Center for Disease Control (CDC)
- Ask a variety of questions related to health and wellness
- Cleaned data with state aggregated values posted on website

## BRFSS Data Aggregated by State

```{r}
states.stats <- read.csv("http://heike.github.io/rwrks/02-r-graphics/data/states.stats.csv")
head(states.stats)
```

## We must join this data again

```{r}
states.map <- left_join(states, states.stats, by = c("region" = "state.name"))
head(states.map)
```

## Numerical Information Using Shade and Intensity

Average number of days in the last 30 days of insufficient sleep by state

```{r, fig.height=4, fig.width=7}
qplot(long, lat, geom = "polygon", data = states.map, 
      group = group, fill = avg.qlrest2) + coord_map()
```

## BRFSS Data Aggregated by State

```{r}
states.sex.stats <- read.csv("http://heike.github.io/rwrks/02-r-graphics/data/states.sex.stats.csv")
head(states.sex.stats)
```

## One More Join

```{r}
states.sex.map <- left_join(states, states.sex.stats, by = c("region" = "state.name"))
head(states.sex.map)
```

## Adding Numerical Information

Average number of alcoholic drinks per day by state and gender

```{r, fig.height=4, fig.width=7}
qplot(long, lat, geom = "polygon", data = states.sex.map, 
      group = group, fill = avg.drnk) + coord_map() + 
    facet_grid(sex ~ .)
```
