---
title: "Plotting Map Data"
author: "Eric Hare, Andee Kaplan, Samantha Tyner"
date: "May 16, 2016"
output: revealjs::revealjs_presentation
---

```{r, echo=FALSE}
library(ggplot2)
```

## States Data

To make a map, let's load up the states data and take a look:

```{r}
states <- map_data("state")

head(states)
```

## Basic Map Data

What needs to be in the data set in order to plot a basic map?
  
- Need latitude/longitude points for all map boundaries
- Need to know which boundary group all lat/long points belong
- Need to know the order to connect points within each group

## Data for Building Basic State Map

Our *states* data has all necessary information

## A Basic (Rather Hideous) Map

A bunch of latitude longitude points...

```{r, fig.height=4, fig.width=7}
qplot(long, lat, geom = "point", data = states)
```

## A Bit Better of a Map

... that are connected with lines in a very specific order.

```{r, fig.height=4, fig.width=7}
qplot(long, lat, geom = "path", data = states, group = group) + 
    coord_map()
```

## Polygon instead of Path

```{r, fig.height=4, fig.width=7}
qplot(long, lat, geom = "polygon", data = states, group = group) + 
    coord_map()
```

## Incorporating Information About States

We want to incorporate additional information into the plot:

- Add other geographic information by adding geometric layers to the plot
- Add non-geopgraphic information by altering the fill color for each state
    - Use `geom = "polygon"` to treat states as solid shapes to add color
    - Incorporate numeric information using color shade or intensity
    - Incorporate categorical informaion using color hue
    
## Categorical Information Using Hue

If a categorical variable is assigned as the fill color then `qplot` will assign different hues for each category. Let's load in a state regions dataset:

```{r}
statereg <- read.csv("http://heike.github.io/rwrks/02-r-graphics/data/statereg.csv")
```


```{r, fig.height=4, fig.width=7}
qplot(long, lat, geom="polygon", data=states.class.map, group=group, fill=StateGroups, colour=I("black")) + coord_map() 
```
