---
title: "Get Going with R"
subtitle: "Iowa State University"
author: Sam Tyner
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
tips <- read.csv("http://heike.github.io/rwrks/01-r-intro/data/tips.csv")
```


## Set Up

### Your background

What do you know already?

- Excel, Numbers, Some sort of spreadsheet software
- SAS, SPSS, Minitab
- Any other programming language
- A text editor
- R

### Install R 

Hopefully this is done already.  If not, for Windows or OS X:

- Go to http://www.r-project.org/
- Click the CRAN link on the left, and pick a download site (0-Cloud is a good choice)
- Choose link based on your OS
- On Windows, choose the "base" subdirectory to install R.
- On OS X, choose the .pkg file to install R.

### Installing RStudio

This should also be done already, but if not...

- Browse to https://www.rstudio.com/
- Click on the "Download RStudio" 
- Download RStudio Desktop Free
- Choose the installer appropriate for your platform

### Let's take a look at R together...

<!--### Exercise 

*Here's a simple exercise with an empty code chunk provided for entering the answer.*

Write the R code required to add two plus two:

```{r two-plus-two, exercise=TRUE}

```

### Exercise with Code

*Here's an exercise with some prepopulated code as well as `exercise.lines = 5` to provide a bit more initial room to work.*

Now write a function that adds any two numbers and then call it:

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```
--> 

## Motivating Example

- Kick off the workshop by exploring a real data set using R!
- Goal: get the flavor of using R for data management and exploration
- Don't worry about understanding all the coding right away
- We will go back and explain how it all works in detail

### Tips Dataset

- Tips data set recorded by a server in a restaurant over a span of about 10 weeks
- Server recorded several variables about groups they served:
    - Amount they were tipped
    - Cost of the total bill
    - Several characteristics about the groups being served
- Primary Question: How do these variable influence the amount being tipped?
<!--Follow along using [2-MotivatingExample.R](../code/2-MotivatingExample.R) --> 

### Load the tips data

First, we use R to look at the top few rows of the our tip data set. First, we load the data using `read.csv`, then use the `head` function to look at the first 6 rows of the data:

```{r read-tips-ex, exercise=TRUE}
tips <- read.csv("http://heike.github.io/rwrks/01-r-intro/data/tips.csv")
head(tips)
```

### Explore the tips data

How big is this data set and what types of variables are in each column? Let's also get a summary of the values for each variable in tips:

```{r explore-tips, exercise=TRUE}
glimpse(tips)
summary(tips)
```

### Plotting the tips data

Lets look at the relationship between total bill and tip value. First, we need to load the `ggplot2` package (more on this later), then we can use the functions in the `ggplot2` package to make a scatterplot of total bill by tip value. 

```{r scatter-tips, exercise=TRUE, fig.align='center'}
library(ggplot2)
ggplot(data = tips, aes(x = total_bill, y = tip)) + 
  geom_point()
```

Let's add some color! Change the color of the points according to the time they ate (lunch or dinner groups).

```{r scatter-tips-col, exercise=TRUE, fig.align='center'}
ggplot(data = tips, aes(x = total_bill, y = tip)) + 
  geom_point(aes(color = time))
``` 

We can also draw the best-fit line (also known as a linear regression line) on the plot: 

```{r, scatter-tips-lm, exercise=TRUE, fig.align='center'}
ggplot(data = tips, aes(x = total_bill, y = tip)) + 
  geom_point(aes(color = time)) + 
  geom_smooth(method = "lm", color = 'green', se = FALSE)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```

## Basics

## Data Structures

## Packages & Basic Programming

## R Markdown 