---
title: "Intro to dplyr"
author: "Eric Hare, Andee Kaplan, Carson Sievert"
date: "June 10, 2015"
output:
  ioslides_presentation: default
---

```{r, echo=FALSE}
library(knitr)
opts_chunk$set(message = FALSE, warning = FALSE)
```

##Baseball Data

 - The `plyr` package contains the data set `baseball`
 - seasonal batting statistics of all major league players (through 2007)
 
 
```{r}
data(baseball, package = "plyr")
help(baseball, package = "plyr")
head(baseball)
```


##Baseball Data
 - We would like to create career summary statistics for each player
 - Plan: subset on a player, and compute statistics

```{r subsets}
ss <- subset(baseball, id=="sosasa01")
head(ss)
mean(ss$h/ss$ab)
```

We need an **automatic** way to calculate this.

##`for` loops
- Idea: repeat the same (set of) statement(s) for each element of an index set
- Setup: 
    - Introduce counter variable (sometimes named `i`)
    - Reserve space for results
- Generic Code:

    ```r
    result <- rep(NA, length(indexset))
    for(i in indexset){
      ... some statments ...
      result[i] <- ...
    }
    ```
    
##`for` loops for Baseball
- Index set: player id
- Setup: 

```{r baseballforloop}
# Index set
players <- unique(baseball$id)
n <- length(players)

# Place to store data
ba <- rep(NA, n)

# Loop
for(i in 1:n){
  career <- subset(baseball, id==players[i])
  ba[i] <- with(career, mean(h/ab, na.rm=T))
}

# Results
summary(ba)
```

##`for` loops for Baseball
- Index set: player id
- `i = 0`: 

```{r baseballforloop2}
# Index set
players <- unique(baseball$id)
n <- length(players)

# Place to store data
ba <- rep(NA, n)

# Results
head(ba)
```

##`for` loops for Baseball
- Index set: player id
- `i = 1`: 

```{r baseballforloop3}
# Index set
players <- unique(baseball$id)

# Place to store data
ba <- rep(NA, length(players))

for(i in 1:1){ #loop
  career <- subset(baseball, id==players[i])
  ba[i] <- with(career, mean(h/ab, na.rm=T))
}

head(ba)
```

##`for` loops for Baseball
- Index set: player id
- `i = 2`: 

```{r baseballforloop4}
# Index set
players <- unique(baseball$id)

# Place to store data
ba <- rep(NA, length(players))

for(i in 1:2){ #loop
  career <- subset(baseball, id==players[i])
  ba[i] <- with(career, mean(h/ab, na.rm=T))
}

head(ba)
```

##YOUR TURN
- MLB rules for the greatest all-time hitters are that players have to have played at least 1000 games with at least as many at-bats in order to be considered
- Extend the for loop above to collect the additional information
- Introduce and collect data for two new variables: `games` and `atbats`

##How did it go? What was difficult?
- household chores (declaring variables, setting values each time) distract from real work
- indices are error-prone
- loops often result in slow code because R can compute quantities using entire vectors in an optimized way

##Summarise
- A special function: `summarise` or `summarize`
```{r baseball-summarize-example}
library(dplyr)
baseball <- read.csv("../data/baseball.csv")
summarise(baseball, ab=mean(h/ab, na.rm=T))

summarise(baseball,
          ba = mean(h/ab, na.rm=T),
          games = sum(g, na.rm=T),
          hr = sum(hr, na.rm=T),
          ab = sum(ab, na.rm=T))

summarise(subset(baseball, id=="sosasa01"), 
          ba = mean(h/ab, na.rm=T),
          games = sum(g, na.rm=T),
          hr = sum(hr, na.rm=T),
          ab = sum(ab, na.rm=T))
```

##`dplyr` + `Summarize`

- A powerful combination to create summary statistics

```{r ddply-summarize-example}
careers <- summarise(group_by(baseball, id),
                 ba = mean(h/ab, na.rm=T),
                 games = sum(g, na.rm=T),
                 homeruns = sum(hr, na.rm=T),
                 atbats = sum(ab, na.rm=T))

head(careers)
```

##Your Turn
- Find some summary statistics for each of the teams (variable `team`)
    - How many different (`unique`) players has the team had?
    - What was the team's first/last season?
- Challenge: Find the number of players on each team over time. Does the number change?


