---
title: "Motivating Example"
author: "Haley Jeppson, Joe Papio,<br>Sam Tyner"
date: "June 12, 2017"
output: 
    revealjs::revealjs_presentation:
        transition: fade
        theme: white
        highlight: zenburn
        css: custom.css

---

## Motivating Example

- Kick off the workshop by exploring a real data set using R!
- Goal: get the flavor of using R for data management and exploration
- Don't worry about understanding all the coding right away
- We will go back and explain how it all works in detail

## Tips Dataset

- Tips data set recorded by a server in a restaurant over a span of about 10 weeks
- Server recorded several variables about groups they served:
    - Amount they were tipped
    - Cost of the total bill
    - Several characteristics about the groups being served
- Primary Question: How do these variable influence the amount being tipped?
- Follow along using [2-MotivatingExample.R](../code/2-MotivatingExample.R)

## First look at data in R

Lets use R to look at the top few rows of the tips data set. First, we load tip using `read.csv`:

```{r}
tips <- read.csv("http://heike.github.io/rwrks/01-r-intro/data/tips.csv")
```

Now, we use the `head` function to look at the first 6 rows of the data:

```{r}
head(tips)
```

## Tips Data Attributes

How big is this data set and what types of variables are in each column?

```{r}
str(tips)
```

## Tips Variables

Let's get a summary of the values for each variable in tips:

```{r}
summary(tips)
```

## Scatterplots

Lets look at the relationship between total bill and tip value. First, we need to install and load `ggplot2`:

```{r, eval=FALSE}
install.packages("ggplot2")
library(ggplot2)
```

```{r, echo=FALSE}
library(ggplot2)
```

```{r, fig.height=3, fig.width=7}
qplot(tip, total_bill, geom = "point", data = tips)
```

## More Scatterplots

Color the points by lunch and dinner groups

```{r, fig.height=4, fig.width=7}
qplot(tip, total_bill, geom = "point", data = tips, colour = time)
```

## Even More Scatterplots

Add linear regression line to the plot

```{r, fig.height=4, fig.width=7}
qplot(tip, total_bill, geom = "point", data = tips) + 
    geom_smooth(method = "lm")
```

## Rate of Tipping

Tipping is generally done using a rule of thumb based on a percentage of the total bill. We will make a new variable in the data set for the tipping rate = tip / total bill

```{r}
tips$rate <- tips$tip / tips$total_bill

summary(tips$rate)
```

## Tipping Rate Histogram

Lets look distribution of tipping rate values with a histogram

```{r, fig.height=4, fig.width=7}
qplot(rate, data = tips, binwidth = .01)
```

## Someone is an AMAZING tipper...

One person tipped over 70%, who are they?

```{r}
tips[which.max(tips$rate),]
```

## Rates by Gender

Look at the average tipping rate for men and women seperately

```{r}
mean(tips$rate[tips$sex == "Male"])
mean(tips$rate[tips$sex == "Female"])
```

## Statistical Significance

There is a difference but is it statistically significant?

```{r}
t.test(rate ~ sex, data = tips)
```

## Boxplots

Perhaps we are interested in if smokers tip at a different rate than non-smokers.  We could compare the rate values of each group with a side by side boxplot!

```{r, fig.height=3, fig.width=7}
qplot(smoker, rate, geom = "boxplot", data = tips)
```

<!--
## Boxplots Again

What about the difference between male and female smokers and non-smokers?

```{r, fig.height=3, fig.width=7}
qplot(smoker, rate, geom = "boxplot", data = tips, color = sex)
```
--> 

## Your Turn {data-background=#527a7a}

Try playing with chunks of code from this session to further investigate the tips data:

1. Get a summary of the total bill values
2. Make side by side boxplots of tip rates for different days of the week
3. Find the average tip value for smokers
