---
title: "Data Structures"
author: "Eric Hare, Andee Kaplan, Sam Tyner"
date: "May 16, 2016"
output: 
    revealjs::revealjs_presentation:
        transition: fade
        theme: white
        highlight: zenburn
---

## Data Frames

- Data Frames are the work horse of R objects
- Structured by rows and columns and can be indexed
- Each column is a specified variable type
- Columns names can be used to index a variable
- Advice for naming variable applys to editing columns names
- Can be specified by grouping vectors of equal length as columns

## Data Frame Indexing

- Elements indexed similar to a vector using `[` `]`
- `df[i,j]` will select the element in the $i^{th}$ row and $j^{th}$ column
- `df[ ,j]` will select the entire $j^{th}$ column and treat it as a vector
- `df[i ,]` will select the entire $i^{th}$ row and treat it as a vector
- Logical vectors can be used in place of i and j used to subset the row and columns

## Adding a new Variable to a Data Frame

- Create a new vector that is the same length as other columns
- Append new column to the data frame using the `$` operator
- The new data frame column will adopt the name of the vector

## Data Frame Demo

Use Edgar Anderson's Iris Data:

```{r}
flower <- iris
```

Select Species column (5th column):

```{r}
flower[,5]
```

## Demo (Continued)

Select Species column with the `$` operator:

```{r}
flower$Species
```

## Demo (Continued)

```{r}
flower$Species == "setosa"
```

## Demo (Continued)

```{r}
flower[flower$Species=="setosa", ]
```

## Creating our own Data Frame

Create our own data frame using `data.frame` function

```{r}
mydf <- data.frame(NUMS = 1:5, 
                   lets = letters[1:5],
                   vehicle = c("car", "boat", "car", "car", "boat"))
mydf
```

## Renaming columns

We can use the `names` function to set that first column to lowercase:

```{r}
names(mydf)[1] <- "nums"
mydf
```

## Your Turn {data-background=#527a7a}

1. Make a data frame with column 1: 1,2,3,4,5,6 and column 2: a,b,a,b,a,b
2. Select only rows with value "a" in column 2 using logical vector
3. `mtcars` is a built in data set like `iris`: Extract the 4th row of the `mtcars` data.

## Lists

- Lists are a structured collection of R objects
- R objects in a list need not be the same type
- Create lists using the `list` function
- Lists indexed using double square brackets `[[ ]]` to select an object

## List Example

Creating a list containing a vector and a matrix:

```{r}
mylist <- list(matrix(letters[1:10], nrow = 2, ncol = 5),
               seq(0, 49, by = 7))
mylist
```

Use indexing to select the second list element:

```{r}
mylist[[2]]
```

## Your Turn {data-background=#527a7a}

1. Create a list containing a vector and a 2x3 data frame
2. Use indexing to select the data frame from your list
3. Use further indexing to select the first row from the data frame in your list

## Examining Objects

- `head(x)` - View top 6 rows of a data frame
- `tail(x)` - View bottom 6 rows of a data frame
- `summary(x)` - Summary statistics   
- `str(x)` - View structure of object  
- `dim(x)` - View dimensions of object 
- `length(x)` - Returns the length of a vector

## Examining Objects Example

We can examine the first two values of an object by passing the `n` parameter to the `head` function:

```{r}
head(iris, n = 2)
```

What's its structure?

```{r}
str(iris)
```

## Your Turn {data-background=#527a7a}

1. View the top 8 rows of mtcars data
2. What type of object is the mtcars data set?
3. How many rows are in iris data set? (try finding this using dim or indexing + length)
4. Summarize the values in each column in iris data set

## Working with Output from a Function

- Can save output from a function as an object
- Object is generally a list of output objects
- Can pull off items from the output for further computing
- Examine object using functions like `str(x)`

## Saving Output Demo

- t-test using iris data to see if petal lengths for setosa and versicolor are the same
- t.test function can only handle two groups, so we subset out the virginica species

```{r}
t.test(Petal.Length ~ Species, data = iris[iris$Species != "virginica", ])
```

## Demo (Continued)

Save the output of the t-test to an object

```{r}
tout <- t.test(Petal.Length ~ Species, data = iris[iris$Species != "virginica", ])
```

Let's look at the structure of this object:

```{r}
str(tout)
```

## Demo: Extracting the P-Value

Since this is simply a list, we can use our regular indexing:

```{r}
tout$p.value
tout[[3]]
```

## Importing Data

We often need to import in our own data rather than just using built-in datasets.

- First need to tell R where the data is saved using `setwd()`
- Data read in using R functions such as:
    - `read.table()` for reading in .txt files
    - `read.csv()` for reading in .csv files
- Assign the data to new R object when reading in the file

## Importing Data Demo

We first create a csv file. We can use a text editor, excel... Then we load it in:

```{r, eval=FALSE}
littledata <- read.csv("PretendData.csv")
```

## Your Turn {data-background=#527a7a}

- Make 5 rows of data in an excel spreadsheet and save it as a *tab-delimited txt file*.
- Import this new .txt file into R with `read.table`. You may need to look at the help page for `read.table` in order to properly do this.
