---
title: "Basic Plots"
author: "Haley Jeppson, Sam Tyner"
output: 
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r, echo=FALSE}
library(ggplot2)
```

## Let's Run the Setup File...

Run this line in the R console. You should see a plot appear if setup is successful.

```{r eval = FALSE}
source("../code/0-Setup.R")
```

---

## ggplot2 In a Nutshell

- Package for statistical graphics
- Developed by Hadley Wickham (An ISU Alumnus)
- Designed to adhere to good graphical practices
- Supports a wide variety plot types
- Constructs plots using the concept of layers\medskip
- http://ggplot2.tidyverse.org/ or Hadley's book *ggplot2: Elegant Graphics for Data Analysis}* for reference material

---

## *qplot* Function

The *qplot()* function stands for "quick plot"

- Produces all plot types available with ggplot2
- Most arguments have defaults unless you specify otherwise
- Allows for plotting options within the function statement
- Creates an object that can be saved
- Plot layers can be added to modify plot complexity
- Resembles *base* plotting more than the rest of ggplot2

---

## *qplot* Structure

The *qplot()* function has a basic syntax:

*qplot(variables, plot type, dataset, options)*

- *variables*: list of variables used for the plot (need this one)

- *plot type*: specified with a *geom =* statement

- *dataset*: specified with a *data =* statement (need this one)

- *options*: there are so, so many options!

---

## Diamonds Data

We will explore the diamonds data set (preloaded along with ggplot2) using qplot for basic plotting.

The data set was scraped from a diamond exchange company data base by Hadley.  It contains the prices and attributes of over 50,000 diamonds

---

## Examining the Diamonds Data

What does the data look like?

Lets look at the top few rows of the diamond data frame to find out!

```{r}
head(diamonds)
```

---

## Basic Scatterplot

Basic scatter plot of diamond price vs carat weight

```{r, fig.height=4, fig.width=7}
qplot(carat, price, geom = "point", data = diamonds)
```

---

## Another Scatterplot

Scatter plot of diamond price vs carat weight showing versitility of options in qplot

```{r, fig.height=3, fig.width=7}
qplot(carat, log(price), geom = "point", data = diamonds, 
	alpha = I(0.2), colour = color, 
	main = "Log price by carat weight, grouped by color") + 
    xlab("Carat Weight") + ylab("Log Price")
```

---
class: inverse

## Your Turn 

All of the your turns for this section will use the tips data set:

```{r}
tips <- read.csv("http://heike.github.io/rwrks/02-r-graphics/data/tips.csv")
```

1. Use qplot to build a scatterplot of variables tips and total bill
2. Use options within qplot to color points by smokers
3. Clean up axis labels and add main plot title

---

## Basic Histogram

Basic histogram of carat weight

```{r, fig.height=4, fig.width=7}
qplot(carat, geom = "histogram", data = diamonds)
```

---

## Another Histogram

Carat weight histograms faceted by cut

```{r, fig.height=4, fig.width=7}
qplot(carat, geom = "histogram", data = diamonds, binwidth = 0.2, facets = .~cut)
```

---
class: inverse
## Your Turn 

1. Create a new variable in tips data frame rate = tip / total bill
2. Use qplot to create a histogram of rate
3. Change the bin width on that histogram to 0.05
4. Facet this histogram by size of the group

---

## Basic Boxplot

Side by side boxplot of diamond prices within cut groupings

```{r, fig.height=4, fig.width=7}
qplot(cut, price, geom = "boxplot", data = diamonds)
```

---

## Another Boxplot

Side by side boxplot of log prices within cut groupings with jittered values overlay

```{r, fig.height=3, fig.width=7}
qplot(cut, log(price), geom = "boxplot", data = diamonds, 
	main = "Boxplots of log Diamond Prices Grouped by Cut Quality") +
	geom_jitter(alpha = I(.025))
```

---
class: inverse

## Your Turn 

1. Make side by side boxplots of tipping rate for males and females
2. Overlay jittered points for observed values onto this boxplot

---

## Bar Plots

To investigate bar plots we will switch over to the Titanic data set:

```{r}
titanic <- as.data.frame(Titanic)
```

Data includes passenger characteristics and survival outcomes for those aboard the RMS Titanic's ill fated maiden voyage

---

## Basic Bar Plot

Basic bar plot of survival outcomes

```{r, fig.height=4, fig.width=7}
qplot(Survived, geom = "bar", data = titanic, weight = Freq)
```

---

## Another Bar Plot

Bar plot faceted by gender and class

```{r, fig.height=4, fig.width=7}
qplot(Survived, geom = "bar", data = titanic, weight = Freq, 
      facets = Sex~Class)
```

---
class: inverse

## Your Turn 

1. Use the tips data to make a barplot for counts of smoking and non smoking customers
2. Facet using day of week and time of day to view how smoking status changes for different meal times
