---
title: "Plotting Map Data"
author: "Haley Jeppson, Joe Papio,<br>Sam Tyner"
date: "June 13, 2017"
output: 
    revealjs::revealjs_presentation:
        transition: fade
        theme: white
        highlight: zenburn
        css: custom.css
---

## Loading Required Packages

```{r, eval=TRUE, message = FALSE, warning=FALSE}
library(dplyr)
library(albersusa)
library(sf)
```

## States Data

To make a map, let's load up the states data and take a look:

```{r alt1, eval = FALSE, echo=FALSE}
library(ggplot2)
states <- map_data("state")
head(states)
```

```{r}
library(ggplot2)
library(albersusa)
library(sf)
states <- usa_sf("laea")
glimpse(states)
```

## Basic Map Data

What needs to be in the data set in order to plot a basic map?
  
- Need latitude/longitude points for all map boundaries
- Need to know which boundary group all lat/long points belong
- Need to know the order to connect points within each group

## Simple Features

- A *feature* is a real-world object - can consist of other objects.
    * e.g. a car is an object that consists of tires, engine, etc.
- Features have *geometry* - where the object is located and description of other properties
- Map data are MULTIPOLYGON simple feature objects
- Borders of states, countries, etc. form POLYGON objects
    * e.g. each US state has a geometry that describes where in the world it is, and various other descriptors like population, land size, etc.

See `vignette("sf1", package = "sf")` for more

<!--
## Data for Building Basic State Map

Our *states* data has all necessary information

## A Basic (Rather Hideous) Map

A bunch of latitude longitude points...

```{r, eval = FALSE,fig.height=4, fig.width=7}
qplot(long, lat, geom = "point", data = states)
```

## A Bit Better of a Map

... that are connected with lines in a very specific order.

```{r, eval = FALSE, fig.height=4, fig.width=7}
qplot(long, lat, geom = "path", data = states, group = group) + 
    coord_map()
```
-->

## Plotting Simple Features

Very easy with `geom_sf()`

```{r, fig.height=4, fig.width=7}
ggplot(data = states) + geom_sf()
```

## Adding State Attributes

We want to incorporate additional information into the plot:

- Add other geographic information by adding geometric layers to the plot
- Add non-geographic information by altering the fill color for each state
    - Use `fill` to treat states as solid shapes to add color
    - Incorporate numeric information using color shade or intensity
    - Incorporate categorical informaion using color hue
    
## Categorical Information Using Hue

If a categorical variable is assigned as the fill color then `ggplot2` will assign different hues for each category. Let's load in a state regions dataset:

```{r}
statereg <- read.csv("http://heike.github.io/rwrks/02-r-graphics/data/statereg.csv", stringsAsFactors = FALSE)

glimpse(statereg)
```

## Joining Data

We need to `join` or `merge` our original states data with this new state info. We can use the `left_join` function to do so (more on this later):

```{r}
states$name <- tolower(states$name)
states.class.map <- left_join(states, statereg, by = c("name" = "State"))
glimpse(states.class.map)
```

## Plotting the Result

```{r alt2, eval = FALSE, echo = FALSE, fig.height=4, fig.width=7}
qplot(long, lat, geom = "polygon", data = states.class.map, 
      group = group, fill = StateGroups, colour = I("black")) + 
    coord_map() 
```

```{r, fig.height=4, fig.width=7}
ggplot(data = states.class.map) + geom_sf(aes(fill = StateGroups))
```

## Numerical Information Using Shade and Intensity

To show how was can add numerical information to map plots we will use the BRFSS data

- Behavioral Risk Factor Surveillance System
- 2008 telephone survey run by the Center for Disease Control (CDC)
- Ask a variety of questions related to health and wellness
- Cleaned data with state aggregated values posted on website

## BRFSS Data Aggregated by State

```{r}
states.stats <- read.csv("http://heike.github.io/rwrks/02-r-graphics/data/states.stats.csv", stringsAsFactors = FALSE)
glimpse(states.stats)
```

## We must join this data again

```{r}
states.map <- left_join(states, states.stats, by = c("name" = "state.name"))
glimpse(states.map)
```

## Shade and Intensity

Average number of days in the last 30 days of insufficient sleep by state

```{r alt3, echo = FALSE, eval = FALSE, fig.height=3, fig.width=7}
qplot(long, lat, geom = "polygon", data = states.map, 
      group = group, fill = avg.qlrest2) + coord_map()
```

```{r, fig.height=3, fig.width=7}
ggplot(data = states.map) + geom_sf(aes(fill = avg.qlrest2))
```

## BRFSS Data Aggregated by State

```{r}
states.sex.stats <- read.csv("http://heike.github.io/rwrks/02-r-graphics/data/states.sex.stats.csv", stringsAsFactors = FALSE)
glimpse(states.sex.stats)
```

## One More Join

```{r}
states.sex.map <- left_join(states, states.sex.stats, by = c("name" = "state.name"))
glimpse(states.sex.map)
```

## Adding Information

Average number of alcoholic drinks per day by state and gender

```{r alt4, echo = FALSE, eval = FALSE, fig.height=3, fig.width=7}
qplot(long, lat, geom = "polygon", data = states.sex.map, 
      group = group, fill = avg.drnk) + coord_map() + 
    facet_grid(sex ~ .)
```

```{r, fig.height=3, fig.width=7}
ggplot(data = states.sex.map) + geom_sf(aes(fill = avg.drnk)) +
  facet_grid(.~sex)
```

## Your Turn {data-background=#527a7a}

- Use `left_join` to combine child healthcare data with maps information. You can load in the child healthcare data with:

```{r}
states.health.stats <- read.csv("http://heike.github.io/rwrks/02-r-graphics/data/states.health.stats.csv", stringsAsFactors = FALSE)
```

- Use `geom_sf` to create a map of child healthcare undercoverage rate by state

## Cleaning Up Your Maps

Use ggplot2 options to clean up your map!

- Adding Titles `+ ggtitle(...)`
- Might want a plain white background `+ theme_bw()`
- Extremely familiar geography may eliminate need for latitude and longitude axes `+ theme(...)`
- Want to customize color gradient `+ scale_fill_gradient2(...)`
- Keep aspect ratios correct `+ coord_map()`

## Cleaned Up Map

```{r alt5, eval = FALSE, echo= FALSE, fig.height=3, fig.width=7}
qplot(long, lat, geom="polygon", data = states.map, group = group, fill = avg.drnk) + 
  coord_map()
```

```{r, fig.height=3, fig.width=7}
ggplot(data = states.map) + geom_sf(aes(fill = avg.drnk)) +
  theme_bw() +
  scale_fill_gradient2(limits = c(1.5, 3),low = "lightgray", high = "red") + 
  theme(axis.ticks = element_blank(),
       axis.text.x = element_blank(),
       axis.title.x = element_blank(),
       axis.text.y = element_blank(),
       axis.title.y = element_blank()) +
  ggtitle("Map of Average Number of Alcoholic Beverages\nConsumed Per Day by State")
```

## Your Turn {data-background=#527a7a}

Use options to polish the look of your map of child healthcare undercoverage rate by state!
