---
title: "Polishing your Plots"
author: "Haley Jeppson, Joe Papio,<br>Sam Tyner"
date: "June 13, 2017"
output: 
  revealjs::revealjs_presentation:
        transition: fade
        theme: white
        highlight: zenburn
        css: custom.css
---

```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(plotly)
library(sf)
library(albersusa)

states <- map_data("state")
```

## Visual Appearance

So far have mostly discussed how to get the data displayed the way you want, focusing on the
essence of the plot.

Themes give you a huge amount of control over the appearance of the plot, the choice of background colours, fonts and so on.

Bonus: Can make plots *interactive* with `plotly`

## Built-In Themes

```{r, fig.height=2, fig.width=7}
ggplot(data = diamonds, aes(x = carat, y = price)) + 
  geom_point()
myplot <- ggplot(data = diamonds, aes(x = carat, y = price)) + 
  geom_point()
```

```{r, fig.height=2, fig.width=7}
myplot + theme_bw()
```

## Setting Themes

You can set a theme for all plots:

```{r, eval=FALSE}
theme_set(theme_bw())
```

We can also take a look at options for each theme:

```{r}
theme_bw()
```

## Plot Title

You can change this for an individual plot

```{r, fig.height=4, fig.width=7}
myplot +
    ggtitle("Price vs Carat for Diamonds")
```

## Elements

You can make your own theme, or modify an existing one.

Themes are made up of elements which can be one of:

- element_line
- element_text
- element_rect
- element_blank

This gives you a lot of control over plot appearance.

## Elements we can set

- **Axis**: axis.line, axis.text.x, axis.text.y, axis.ticks, axis.title.x, axis.title.y
- **Legend**: legend.background, legend.key, legend.text, legend.title
- **Panel**: panel.background, panel.border, panel.grid.major, panel.grid.minor
- **Strip**: strip.background, strip.text.x, strip.text.y

## Modifying a plot

```{r, fig.height=4, fig.width=7}
my2ndplot <- myplot + 
    ggtitle("Price vs Carat for Diamonds")
my2ndplot + theme(plot.title = element_text(colour = "red"), 
          axis.text.y = element_text(colour = "purple", angle=45))
```

## Removing Axes

We could also choose to remove all axes (helpful for maps):

```{r, fig.height=3, fig.width=7}
my2ndplot + theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks.length = unit(0, "cm")
)
```

## Saving your Work

The `ggsave()` function will automatically save the last plot produced:

```{r, eval=FALSE}
ggplot(data = diamonds, aes(x = carat, y = price)) + 
  geom_point()

ggsave("diamonds.png")
ggsave("diamonds.pdf")
ggsave("diamonds.png", width = 6, height = 6)
```

We can also explicitly tell it which plot to save:

```{r, eval=FALSE}
ggsave("diamonds.png", plot = myplot, dpi = 72)
```


## Your Turn {data-background=#527a7a}

1. Save a pdf of a scatterplot of price vs carat
2. Open up the pdf in Adobe Acrobat (or another PDF Reader)
3. Save a png of the same scatterplot

## Making plots interactive!

```{r eval = FALSE}
# devtools::install_github("ropensci/plotly")
library(plotly)
dplot <- qplot(carat, price, data = diamonds[sample(53940, 1000),])
ggplotly(dplot)
```

<iframe src="images/plotly_ex.html" style="width:500px;height:500px"></iframe>

